Geometric variational problems

### 2020-11-13 NSF2020/Description/newspecificaim2.tex
yny addition

\subsection{Specific Aim 3: Efficient high-order numerical methods for
large-scale simulations}
\label{subsec:specific_aim_3}
% ----------------------------------------------------------------------
The HAP model requires solving exterior Dirichlet problems for the
screened Laplace equation and the Stokes mobility problem at each time
step in complex domains such as in Figure~\ref{fig:domain}. If
discretizing these equations with stencil-based numerical methods, the
computational domain must be truncated, the fluid volume must be
discretized, and artificial boundary conditions must be imposed. The
artificial boundary conditions introduce additional error, and
discretizing the volume results in an excessively large linear system.
Moreover, it is difficult to obtain high-order discretizations when the
boundary surfaces are irregular. Alternatively, since we are solving
constant-coefficient elliptic PDEs, the solution of the PDEs can be
represented as layer potentials
\begin{align}
  \label{eq:LP}
  u(\mathbf{x}) = \int_{\partial\Omega} K(\mathbf{x},\mathbf{y})
  \sigma(\mathbf{y})\,dS,
\end{align}
where $\sigma$ is an unknown density function, and
$K(\mathbf{x},\mathbf{y})$ is a derivative of the fundamental solution
of the underlying PDE. By matching the boundary condition at
$\mathbf{x}_0 \in \partial\Omega$ with the limit of
equation~\eqref{eq:LP} as $\mathbf{x}\rightarrow \mathbf{x}_0$, a BIE
for the density function is formed. Numerical methods to solve BIEs have
several advantages over their PDE-based counterpart: by construction,
the layer potential~\eqref{eq:LP} satisfies far-field conditions; only
$\partial\Omega$ needs to be discretized; carefully chosen kernels
$K(\mathbf{x},\mathbf{y})$ result in a well-conditioned linear system
that is solved with a mesh-independent number of GMRES iterations; and
high-order or spectral accuracy is attainable with appropriate
quadrature methods. In our previous work~\cite{Fu2018_SIAM} we
demonstrate that BIEs are a powerful tool to simulate two-dimensional
suspensions of amphiphilic particles. In addition to improving
two-dimensional simulations, we will extend the results to three
dimensions using a standard three-dimensional BIE formulation of the
screened Laplace equation~\cite{ying_2006} and a well-conditioned BIE
formulation of the mobility problem~\cite{manasthesis, rac-gre2016}.

\begin{wrapfigure}[21]{r}{2.5in}
%\vspace*{+15pt}
\centerline{\includegraphics[width=2.5in]{figures/integrands}}
\vspace*{-13pt}
  \caption{{\footnotesize The Fourier modes of
  $K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ where
  $|\mathbf{x_0}| = 0.99$ and $\mathbf{y} \in \partial\Omega$, and
  $\Omega$ is the unit circle. The trapezoid rule applied to the red
  integrand has large amounts of error, but the trapezoid rule is much
  more accurate when applied to the blue integrand. Since the difference
  between the red and blue curves can be accurately computed with the
  barycentric quadrature rule, the quadrature error applied to layer
  potentials for the screen Laplace equation will be uniformly bounded
  in $\Omega$.}}
\label{fig:integrands}
\end{wrapfigure}
Significant challenges to solve BIEs include solving dense linear
systems, developing preconditioning strategies, and developing
quadrature methods for integrands that are nearly singular. These
challenges are described in \S\ref{subsec:NumericalIssues}. However,
numerical errors can still lead to unphysical contact between bodies. We
propose two algorithms to avoid contact in \S\ref{subsec:timeStepping}.
Then, proposed work for well-conditioned BIE formulations of fluctuating
hydrodynamics are described in \S\ref{subsec:fluctuating}.

@@

Nearly touching bodies is ubiquitous in self-assembly of amphiphilic
particles, and this results in nearly-singular integrands. Quadrature
methods for these integrands has received a lot of attention in two and
three dimensions~\cite{alpert, kapur, sidi, duffy, bruno1, bruno2,
davis_1984, graglia_2008, hackbusch_sauter_1994, jarvenpaa_2003,
khayat_2005, schwab_1992, ying_2006, beale1, beale2, goodman_1990,
haroldson_1998, lowengrub_1993, schwab_1992, ggq1, ggq2, ggq3,
helsing_2008a, helsing_integral_2009, helsing_tutorial_2012,
klockner2013jcp, qbx2, wala2019jcp, af2018sisc, siegel2018jcp,
rachh2017jcp, ding2019arxiv, bar2014}. The trapezoid rule is the
workhorse for two-dimensional BIEs since it achieves spectral accuracy
when integrands are not nearly-singular~\cite{tre-wei2014}. To address
nearly-singular integrands of two-dimensional BIEs, we will use a {\em
barycentric quadrature rule} that only requires a slight modification of
the trapezoid rule~\cite{ioa-pap-per1991}. In its current form, this
method requires the layer potential to satisfy Laplace or Stokes
equations~\cite{bar-wu-vee2015, chi-moo-qua2020}, but the PIs will extend
the method to layer potentials for the screened Laplace
equation. This will be done by recognizing that the fundamental solution
of the screened Laplace equation can be decomposed as
$K(\mathbf{x},\mathbf{y}) = K_1(\mathbf{x},\mathbf{y}) +
K_2(\mathbf{x},\mathbf{y})$, where $K_1(\mathbf{x},\mathbf{y}) =
-\log|\mathbf{x} - \mathbf{y}|$ and $K_2(\mathbf{x},\mathbf{y}) =
K(\mathbf{x},\mathbf{y}) + \log|\mathbf{x} - \mathbf{y}|$. Using this
decomposition, layer potentials involving $K_1(\mathbf{x},\mathbf{y})$
can be accurately computed with the barycentric quadrature
rule~\cite{ioa-pap-per1991}, and layer potentials involving
$K_2(\mathbf{x},\mathbf{y})$ can be accurately computed with the
trapezoid rule since the kernel is bounded for all $\mathbf{x}$ and
$\mathbf{y}$. The proposed method is demonstrated in
Figure~\ref{fig:integrands} where the amplitude of the Fourier modes of
$K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ are
plotted. The quadrature error of the trapezoid rule is the sum of the
unresolved Fourier modes.

@@

Our previous work avoids contact by including a Leonard-Jones body
forces~\cite{Fu2018_SIAM}. Such a steep steric interaction at short
ranges introduces numerical stiffness and limits the time step size. To
remove this stiffness, we propose a geometry-based contact
method~\cite{har-pon-sor-zor2011}, and this method has been applied to
vesicle suspensions~\cite{lu-rah-zor2017} and rigid body suspensions in
two dimensions~\cite{bys-sha-qua2020} and three
dimensions~\cite{Yan2019}. These methods determine repulsion force by
solving a non-linear complementarity problem with a geometric constraint
that the configuration is non-overlapping.

>>>>

\subsection{Specific Aim 2: Efficient high-order numerical methods for
large-scale simulations}
\label{subsec:specific_aim_2}
%\begin{figure}[h]
\begin{wrapfigure}[]{r}{0.60\textwidth}
  \centering{\includegraphics[width=0.6\textwidth]{Figures/domain.pdf}}
    \caption{\label{fig:flow_map} \todo[inline]{Missing caption}}
\end{wrapfigure}
%\end{figure}
% ----------------------------------------------------------------------
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces, van der Waals
interactions, and thermal noise~\cite{Zhang2017}, to coordinate movement
and bind nano-legos in a solvent \cite{Pandey2011, Leong2007, Reynolds2019,
Dasgupta2017, Siontorou2017}. Additionally, hydrodynamic interactions are central for complex
microscopic three-dimensional structures to fabricate~\cite{Cho2010}.
%
In Specific Aim 2, we propose to develop efficient high-order numerical methods to simulate dynamic
assembly of particles under HAP potential in a viscous solvent. 
%
%
%hydrodynamics effects by em
%particles in a solvent. 
%
% Results from these proposed numerical simulations will be useful to 
% Finally, coarse-grained and molecular
%dynamics theories include water either explicitly or implicitly in their
%models. 
%
%The immediate goal of the present proposal is to simulate
%vesicle bilayers in external flows, where the bilayers are represented
%by a self-assembled collection of rigid particles.
%
As described in~\eqref{eq:stokes} and
illustrated in Figure~\ref{fig:flow_map}, the solvent phase is modeled
by the Stokes equations for an incompressible viscous fluid, and these
equations are coupled to the screened-Laplace equation~\eqref{SL}
through viscous and hydrophobic stress balance. 
Hydrodynamic effects can
not be ignored since the rates of biological functions like fusion,
fission, and pore dynamics rely on viscous
dissipation~\cite{RYHAM20112929}. 
%While hydrophobic attraction causes the particles to self-assemble indepedently of the dissipation mechanism, the trajectories and
%equilibrium configurations depend on the fact that the particles travel through a viscous environment.

The HAP model requires solving exterior Dirichlet problems for the
screened Laplace equation and the Stokes mobility problem at each time
step in complex domains such as in Figure~\ref{fig:domain}. If
discretizing these equations with stencil-based numerical methods, the
computational domain must be truncated, the fluid volume must be
discretized, and artificial boundary conditions must be imposed. The
artificial boundary conditions introduce additional error, and
discretizing the volume results in an excessively large linear system.
Moreover, it is difficult to obtain high-order discretizations when the
boundary surfaces are irregular. Alternatively, since we are solving
constant-coefficient elliptic PDEs, the solution of the PDEs can be
represented as layer potentials
\begin{align}
  \label{eq:LP}
  u(\mathbf{x}) = \int_{\partial\Omega} K(\mathbf{x},\mathbf{y})
  \sigma(\mathbf{y})\,dS,
\end{align}
where $\sigma$ is an unknown density function, and
$K(\mathbf{x},\mathbf{y})$ is a derivative of the fundamental solution
of the underlying PDE. By matching the boundary condition at
$\mathbf{x}_0 \in \partial\Omega$ with the limit of
equation~\eqref{eq:LP} as $\mathbf{x}\rightarrow \mathbf{x}_0$, a BIE
for the density function is formed.

Numerical methods to solve BIEs have
several advantages over their PDE-based counterpart: by construction,
the layer potential~\eqref{eq:LP} satisfies far-field conditions; only
$\partial\Omega$ needs to be discretized; carefully chosen kernels
$K(\mathbf{x},\mathbf{y})$ result in a well-conditioned linear system
that is solved with a mesh-independent number of GMRES iterations; and
high-order or spectral accuracy is attainable with appropriate
quadrature methods. In our previous work~\cite{Fu2018_SIAM} we
demonstrate that BIEs are a powerful tool to simulate two-dimensional
suspensions of amphiphilic particles. In addition to improving
two-dimensional simulations, we will extend the results to three
dimensions using a standard three-dimensional BIE formulation of the
screened Laplace equation~\cite{ying_2006} and a well-conditioned BIE
formulation of the mobility problem~\cite{manasthesis, rac-gre2016}.

Significant challenges to solve BIEs include solving dense linear
systems, developing preconditioning strategies, and developing
quadrature methods for integrands that are nearly singular. These
challenges are described in \S\ref{subsec:NumericalIssues}. However,
numerical errors can still lead to unphysical contact between bodies. We
propose two algorithms to avoid contact in \S\ref{subsec:timeStepping}.
Then, proposed work for well-conditioned BIE formulations of fluctuating
hydrodynamics are described in \S\ref{subsec:fluctuating}.

@@

%
\begin{wrapfigure}[21]{r}{2.5in}
%\vspace*{+15pt}
\centerline{\includegraphics[width=2.5in]{figures/integrands}}
\vspace*{-13pt}
  \caption{{\footnotesize The Fourier modes of
  $K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ where
  $|\mathbf{x_0}| = 0.99$ and $\mathbf{y} \in \partial\Omega$, and
  $\Omega$ is the unit circle. The trapezoid rule applied to the red
  integrand has large amounts of error, but the trapezoid rule is much
  more accurate when applied to the blue integrand. Since the difference
  between the red and blue curves can be accurately computed with the
  barycentric quadrature rule, the quadrature error applied to layer
  potentials for the screen Laplace equation will be uniformly bounded
  in $\Omega$.}}
\label{fig:integrands}
\end{wrapfigure}
Nearly touching bodies is ubiquitous in self-assembly of amphiphilic
particles, and this results in nearly-singular integrands. Quadrature
methods for these integrands has received a lot of attention in two and
three dimensions~\cite{alpert, kapur, sidi, duffy, bruno1, bruno2,
davis_1984, graglia_2008, hackbusch_sauter_1994, jarvenpaa_2003,
khayat_2005, schwab_1992, ying_2006, beale1, beale2, goodman_1990,
haroldson_1998, lowengrub_1993, schwab_1992, ggq1, ggq2, ggq3,
helsing_2008a, helsing_integral_2009, helsing_tutorial_2012,
klockner2013jcp, qbx2, wala2019jcp, af2018sisc, siegel2018jcp,
rachh2017jcp, ding2019arxiv, bar2014}. The trapezoid rule is the
workhorse for two-dimensional BIEs since it achieves spectral accuracy
when integrands are not nearly-singular~\cite{tre-wei2014}. To address
nearly-singular integrands of two-dimensional BIEs, we will use a {\em
barycentric quadrature rule} that only requires a slight modification of
the trapezoid rule~\cite{ioa-pap-per1991}. In its current form, this
method requires the layer potential to satisfy Laplace or Stokes
equations~\cite{bar-wu-vee2015, chi-moo-qua2020}, but the PIs will extend
the method to layer potentials for the screened Laplace
equation. This will be done by recognizing that the fundamental solution
of the screened Laplace equation can be decomposed as
$K(\mathbf{x},\mathbf{y}) = K_1(\mathbf{x},\mathbf{y}) +
K_2(\mathbf{x},\mathbf{y})$, where $K_1(\mathbf{x},\mathbf{y}) =
-\log|\mathbf{x} - \mathbf{y}|$ and $K_2(\mathbf{x},\mathbf{y}) =
K(\mathbf{x},\mathbf{y}) + \log|\mathbf{x} - \mathbf{y}|$. Using this
decomposition, layer potentials involving $K_1(\mathbf{x},\mathbf{y})$
can be accurately computed with the barycentric quadrature
rule~\cite{ioa-pap-per1991}, and layer potentials involving
$K_2(\mathbf{x},\mathbf{y})$ can be accurately computed with the
trapezoid rule since the kernel is bounded for all $\mathbf{x}$ and
$\mathbf{y}$. The proposed method is demonstrated in
Figure~\ref{fig:integrands} where the amplitude of the Fourier modes of
$K(\mathbf{x_0},\mathbf{y})$ and $K_2(\mathbf{x_0},\mathbf{y})$ are
plotted. The quadrature error of the trapezoid rule is the sum of the
unresolved Fourier modes.

@@

%
\begin{wrapfigure}[13]{r}{0.30\textwidth}
\centerline{\includegraphics[width=0.30\textwidth]{figures/BIError.pdf}}
  \vspace{-8pt}
\caption{
\label{fig:bierror}  
\footnotesize The false color map shows how numerical quadrature of
  layer potentials loses accuracy near the particle boundaries.  The
  color bar is for $\log_{10}.$}
\end{wrapfigure}
%
Our previous work avoids contact by including a Leonard-Jones body
forces~\cite{Fu2018_SIAM}. Such a steep steric interaction at short
ranges introduces numerical stiffness and limits the time step size. To
remove this stiffness, we propose a geometry-based contact
method~\cite{har-pon-sor-zor2011}, and this method has been applied to
vesicle suspensions~\cite{lu-rah-zor2017} and rigid body suspensions in
two dimensions~\cite{bys-sha-qua2020} and three
dimensions~\cite{Yan2019}. These methods determine repulsion force by
solving a non-linear complementarity problem with a geometric constraint
that the configuration is non-overlapping.

%-------------------------------------------------------------------------------
\subsubsection{Novel reciprocal identities}
The force and torque formulas~\eqref{forceandtorque} require the
hydrophobic stress along the particle boundaries. Unfortunately,
standard quadrature formulas to estimate $\nabla u$ on the particle
boundaries introduces large amounts of quadrature error
(Figure~\ref{fig:bierror}). Therefore, it is useful to devise reciprocal
identities for the force and torque on body $i$ that does not
require integration along body $i$. One identity, that we have proved,
is
\begin{align}
    \label{eq:reciprocal}
{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf
  x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} +
  \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, 
\end{align}
for $i=1,\ldots,N$. Here, $u_i$ is the solution of the screened Laplace
equation when only the contribution from particle $i$ is considered and
$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i
\cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$.
Equation~\eqref{eq:reciprocal} does not require knowledge of $\nabla u$
on the particle boundary, and this is enormously beneficial for
calculating force and torque.

### 2020-11-13 NSF2020/Description/newspecificaim3.tex
yny addition

\subsection{Specific Aim 2: Hydrodynamics}
\label{subsec:specific_aim_2}
In Specific Aim 2, we include hydrodynamics effects by embedding the
particles in a solvent.  As described in~\eqref{eq:stokes} and
illustrated in Figure~\ref{fig:flow_map}, the solvent phase is modeled
by the Stokes equations for an incompressible viscous fluid, and these
equations are coupled to the screened-Laplace equation~\eqref{SL}
through viscous and hydrophobic stress balance. Hydrodynamic effects can
not be ignored since the rates of biological functions like fusion,
fission, and pore dynamics rely on viscous
dissipation~\cite{RYHAM20112929}. 
%While hydrophobic attraction causes the particles to self-assemble indepedently of the dissipation mechanism, the trajectories and
%equilibrium configurations depend on the fact that the particles travel through a viscous environment. 
Additionally, hydrodynamic interactions are central for complex
microscopic three-dimensional structures to fabricate~\cite{Cho2010}.
Over the past decade, there has been an explosion of interest in
small-scale processes that utilize capillary forces, van der Waals
interactions, and thermal noise~\cite{Zhang2017}, to coordinate movement
and bind material subunits~\cite{Pandey2011, Leong2007, Reynolds2019,
Dasgupta2017, Siontorou2017}. Finally, coarse-grained and molecular
dynamics theories include water either explicitly or implicitly in their
models. The immediate goal of the present proposal is to simulate
vesicle bilayers in external flows, where the bilayers are represented
by a self-assembled collection of rigid particles.

%\begin{figure}[h]
\begin{wrapfigure}[]{r}{0.60\textwidth}
  \centering{\includegraphics[width=0.6\textwidth]{Figures/domain.pdf}}
    \caption{\label{fig:flow_map} \todo[inline]{Missing caption}}
\end{wrapfigure}
%\end{figure}

%We must solve the system 
%\setcounter{midequation}{\theequation}
%\addtocounter{midequation}{2}
%\begin{minipage}[t]{0.44\textwidth}
%\begin{subequations}
%\begin{alignat}{2}
%\label{St1} -& \mu \Delta {\bf u} + \nabla p = {\bf 0}, \\
%\label{St2}  & \nabla \cdot {\bf u} = 0,                &&   \text{in } \Omega\\
%\label{St3}  & {\bf u }({\bf x}) = {\bf v}_i + {\bf w_i}\times {\bf x}, \;  && \text{for } {\bf x} \in \partial P_i\\
%\label{St4}  & ({\bf u } - {\bf u}_{\infty})({\bf x}) \to {\bf 0} && \text{as } {\bf x} \to \infty\\
%\notag \\
%\tag{\themidequation}
%\label{HSB1}  & \int_{\partial P_i} \boldsymbol{\sigma} \boldsymbol{\nu} \,\dif S + {\bf F}_{i} = {\bf 0}\span \span
%\end{alignat}
%\end{subequations}
%\end{minipage}
%\addtocounter{equation}{1}
%\setcounter{midequation}{\theequation}
%\addtocounter{midequation}{2}
%\begin{subequations}
%\begin{minipage}[t]{0.5\textwidth}
%\begin{alignat}{2}
%\label{SL1}  - & \rho^2 \Delta u +u = 0, && \text{in } \Omega\\
%\label{SL2}   & u({\bf x}) = f_i({\bf x}),\quad  && \text{for } {\bf x} \in \partial P_i\\
%\label{SL3} &  u({\bf x}) \to 0 && \text{as } {\bf x} \to \infty \\
%\notag \\
%\notag \\
%\tag{\themidequation}
%\label{HSB2}   & \int_{\partial P_i} {\bf x} \times \boldsymbol{\sigma} \boldsymbol{\nu} \,\dif S + {\bf G}_{i} = {\bf 0} \span \span\\
%\notag
%\end{alignat}
%\end{minipage} 
%\end{subequations}
%\noindent for $i = 1,\dots, N$.

@@

@@

\begin{wrapfigure}[13]{r}{0.30\textwidth}
\centerline{\includegraphics[width=0.30\textwidth]{figures/BIError.pdf}}
  \vspace{-8pt}
\caption{
\label{fig:bierror}  
\footnotesize The false color map shows how numerical quadrature of
  layer potentials loses accuracy near the particle boundaries.  The
  color bar is for $\log_{10}.$}
\end{wrapfigure}

@@

\subsubsection{Novel reciprocal identities}
The force and torque formulas~\eqref{forceandtorque} require the
hydrophobic stress along the particle boundaries. Unfortunately,
standard quadrature formulas to estimate $\nabla u$ on the particle
boundaries introduces large amounts of quadrature error
(Figure~\ref{fig:bierror}). Therefore, it is useful to devise reciprocal
identities for the force and torque on body $i$ that does not
require integration along body $i$. One identity, that we have proved,
is
\begin{align}
    \label{eq:reciprocal}
{\bf F}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i}[\boldsymbol{\sigma}_{ij} + \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S,\quad
{\bf G}_{\text{hydro},i} = \sum_{j \neq i} \int_{\partial P_i} ({\bf
  x}-\mathbf{a}_i) \times [\boldsymbol{\sigma}_{ij} +
  \boldsymbol{\sigma}_{ji}]\boldsymbol{\nu}\,\dif S, 
\end{align}
for $i=1,\ldots,N$. Here, $u_i$ is the solution of the screened Laplace
equation when only the contribution from particle $i$ is considered and
$\boldsymbol{\sigma}_{ij} = \rho^{-1} u_iu_j {\bf I} + \rho(\nabla u_i
\cdot \nabla u_j {\bf I} - 2 \nabla u_i \otimes \nabla u_j)$.
Equation~\eqref{eq:reciprocal} does not require knowledge of $\nabla u$
on the particle boundary, and this is enormously beneficial for
calculating force and torque.

>>>>

\subsection{Specific Aim 3: Dynamics of self-assembly by HAP under an external flow/field}
\label{subsec:specific_aim_3}

@@

%\begin{wrapfigure}[13]{r}{0.30\textwidth}
%\centerline{\includegraphics[width=0.30\textwidth]{figures/BIError.pdf}}
%  \vspace{-8pt}
%\caption{
%\label{fig:bierror}  
%\footnotesize The false color map shows how numerical quadrature of
%  layer potentials loses accuracy near the particle boundaries.  The
%  color bar is for $\log_{10}.$}
%\end{wrapfigure}

@@

@@

\subsubsection{Effects of electric fields on the dynamic self-assembly}
